Scalable Graph-Based Neural \\ Reconstruction for Connectomics

### 2018-03-13 papers/eccv2018/0_introduction.tex
papers: added much more on biological constraints

\begin{figure}[t]
	\subfloat[]{
	\includegraphics[width=.47\linewidth]{figures/schema/pre-multicut-with-arrows.pdf}
	}
	\subfloat[]{
	\includegraphics[width=.47\linewidth]{figures/schema/post-multicut.png}
	}
	\caption{Example improvement of neural reconstruction. (a) We extract 3D skeletons from pixel-based segmentation algorithms to create a graph representation. Red arrows highlight errors in the input segmentation. (b) We improve the segmentation accuracy using a graph partitioning algorithm, leveraging both local and global information.}
	\label{fig:teaser}
\end{figure}

@@

We present a \textit{top-down graph-based} method that builds on the outputs of bottom-up pixel-based segmentation approaches. 
We first extract skeletons from labels in the input segmentation and generate a simplified graph (Fig.~\ref{fig:teaser}a). 
Next, a trained CNN classifier detects split errors in the segmentation data. 
Split errors correspond to neurons which are incorrectly segmented with multiple labels. 
This classifier produces probabilities that two segments belong to the same neuron.
These probabilities correspond to edge weights in our graph extracted from the input segmentation.
We then use a graph optimization algorithm to partition the graph into a final improved reconstruction by enforcing domain-specific global constraints from biology (Fig.~\ref{fig:teaser}b).

@@

This work makes the following contributions: (1) a novel top-down framework that extracts graphs from skeletonized 3D networks for improved neural reconstruction of connectomics data; (2) a region-based CNN classifier to detect errors; (3) an empirical evaluation of our method on several connectomics datasets; (4) improved performance over a state-of-the-art pixel-based reconstruction approach on average by \TODO{XX}\% with suitable running times.

>>>>

\begin{figure}[t!]
	\centering
	\includegraphics[width=\linewidth]{./figures/teaser_v2.png}
	\caption{Our proposed framework uses both local and global biological constraints to improve an input segmentation.}
	\label{fig:teaser}
\end{figure}

@@

We present a \textit{top-down graph-based} method that builds on the outputs of bottom-up pixel-based segmentation approaches. 
Our method imposes both local and global biological constraints onto the output segmentation to match the underlying structure of neuronal processes.
We first extract skeletons from labels in the input segmentation and generate a simplified graph (Fig.~\ref{fig:teaser}a). 
Next, we train a CNN classifier to learn biological constraints on the local structural shape of a neuron (Fig.~\ref{fig:teaser}b). 
This classifier produces probabilities that two segments belong to the same neuron based on the input shapes of the segments.
These probabilities correspond to edge weights in our graph extracted from the input segmentation.
We then use a graph optimization algorithm to partition the graph into a final improved reconstruction by enforcing domain-specific topological constraints from biology (Fig.~\ref{fig:teaser}c).

@@

This work makes the following contributions: (1) a novel top-down framework that extracts graphs from skeletonized 3D segmentations for improved neural reconstruction of connectomics data; (2) a region-based CNN classifier to detect errors; (3) an empirical evaluation of our method on several connectomics datasets; (4) improved performance over a state-of-the-art pixel-based reconstruction approach on average by \TODO{XX}\% with suitable running times.

### 2018-03-13 papers/eccv2018/1_related_work.tex
papers: added much more on biological constraints

@@

\subsubsection{Region-based methods}

Some early techniques apply computationally expensive graph partitioning algorithms with a single node per superpixel~\cite{andres2012globally}. 
Several pixel-based approaches generate probabilities that neighboring pixels belong to the same neuron.
Often a watershed algorithm will then cluster pixels into super-pixels~\cite{zlateski2015image}.
Many methods build on top of these region-based strategies and train random-forest classifiers to produce the final segmentations~\cite{seymour2016rhoananet,nunez2014graph,parag2017anisotropic,zlateski2015image,10.1371/journal.pone.0125825}.
Some methods rely solely on the accuracy of the predicted affinities and agglomerate hierarchically based on the mean affinity between supervoxels~\cite{lee2017superhuman,funke2017deep}.
Pape et al. present a scalable multicut algorithm to partition superpixels~\cite{beier2017multicut}.

>>>>

\begin{figure}[t]
	\subfloat[]{
		\includegraphics[width=.47\linewidth]{figures/schema/pre-multicut-with-arrows.pdf}
	}
	\subfloat[]{
		\includegraphics[width=.47\linewidth]{figures/schema/post-multicut.png}
	}
	\caption{Example improvement of neural reconstruction. (a) We extract 3D skeletons from pixel-based segmentation algorithms to create a graph representation. Red arrows highlight errors in the input segmentation. (b) We improve the segmentation accuracy using a graph partitioning algorithm, leveraging both local and global information.}
	\label{fig:improved-reconstruction}
\end{figure}

@@

\subsubsection{Region-based methods}
Some early techniques apply computationally expensive graph partitioning algorithms with a single node per superpixel~\cite{andres2012globally}. 
Several pixel-based approaches generate probabilities that neighboring pixels belong to the same neuron.
Often a watershed algorithm will then cluster pixels into super-pixels~\cite{zlateski2015image}.
Many methods build on top of these region-based strategies and train random-forest classifiers to produce the final segmentations~\cite{seymour2016rhoananet,nunez2014graph,parag2017anisotropic,zlateski2015image,10.1371/journal.pone.0125825}.
Some methods rely solely on the accuracy of the predicted affinities and agglomerate hierarchically based on the mean affinity between supervoxels~\cite{lee2017superhuman,funke2017deep}.
Pape et al. present a scalable multicut algorithm to partition superpixels~\cite{beier2017multicut}.

### 2018-03-13 papers/eccv2018/2_method.tex
eccv: fixed figures to be horizontally aligned
papers: added much more on biological constraints

\begin{figure*}[htbp]
	\begin{center}
		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/multicut_input.png}}
		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/pre-multicut.png}}
		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/multicut-graph.png}}
		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/post-multicut.png}}
	\end{center}
	\caption{Outline of our approach, from left to right: result of the pixel-based segmentation and agglomeration algorithm; segments of several neurons from the initial segmentation; extracted skeletonized network of those segments; improved 3D reconstruction of the selected segments after graph construction and partitioning with constraints.}
	\label{fig:overview}
\end{figure*}

@@

From the input segmentation we generate a graph $G$ with nodes $N$ and edges $E$ with weights $w_e$. 
The nodes correspond to labeled segments from the segmentation.
There are edges between all nodes that we consider as merge candidates.
Ideally, our graph has edges corresponding to all of the segments that were erroneously split. 
To compute this graph we generate a skeleton for every segment in the pixel-based segmentation (Fig.~\ref{fig:overview}). 
The skeletonized 3D network is a simplified representation of the overall branching structure of the neurons. 
From these skeletons we identify potential merge locations and produce the corresponding edges for the graph. 
To find actual merges we run a classification CNN to generate edge weights corresponding to merge probabilities.
We formulate our graph partitioning problem as a multicut problem to segment the nodes into correct neuronal processes. 
We will now discuss the three major components to our framework (graph creation, edge weight assignment, and graph partitioning) in more detail.

\subsection{Node Generation}
%\subsubsection{Node Generation}
\label{sec:skeletonization}

The simplest node generation strategy creates one node for every unique segment label in the input volume. 
However, some of the labels in the volume correspond to very small structures that are likely the result of segmentation errors, typically in regions with noisy raw image data. 
It is difficult to extract useful shape features from these segments because of their small, often random, shape. 
We prune these nodes from the graph by removing all segments with fewer than a threshold $t_{seg} = 20,000$ voxels. 
This removes on average \TODO{XX}\% of the segments in our  datasets (Sec.~\ref{sec:dataset}), leaving us with around $\TODO{XX}$ nodes per dataset (\TODO{X nodes per micrometer}). 
Despite the large number of segments, these regions only take up \TODO{XX}\% of the total volume on average.
We provide further analysis of this strategy in the supplemental material.

\subsection{Edge Generation}

\begin{figure}[t]
	\centering
	\includegraphics[width=0.45\linewidth]{./figures/skeleton1.png}
	\includegraphics[width=0.45\linewidth]{./figures/skeleton2.png}
	\includegraphics[width=0.45\linewidth]{./figures/skeleton3.png}
	\includegraphics[width=0.45\linewidth]{./figures/skeleton4.png}
	\caption{Example skeletons (in black) extracted from segments using a variant of the TEASER algorithm.}
	\label{fig:skeletonization}
\end{figure}

@@

First, we extract a skeleton from each segment in the label volume using a variant~\cite{zhao2014automatic} of the TEASER algorithm~\cite{sato2000teasar}. 
This skeletonization algorithm repeatedly uses Dijkstra's algorithm to find the farthest voxel from a seed location. 
Since this algorithm is non-linear in the number of voxels, we downsample the datasets so that there are voxel samples every $\SI{30}{\nano\meter}$ in each dimension.
Empirically, this reduced the running time for skeletonization by $\TODO{XX\%}$, with minimal reduction in skeleton accuracy ($\TODO{XX\%}$ fewer branches). 
Fig.~\ref{fig:skeletonization} shows three examples of extracted skeletons (in black). 
These skeletons consist of a sequence of \textit{joints}, i.e., locations that are locally a maximum distance from the segment boundary, with line segments connecting successive joints. 
We refer to joints that have only one connected neighbor as \textit{endpoints}. 
We find that approximately \TODO{XX\%} of the segments that are erroneously split have nearby endpoints  (Fig.~\ref{fig:merge_candidates}). 
We make use of this fact to find merge candidates with the following two-pass pruning algorithm.

In the first pass, we iterate over all endpoints $e$ belonging to a segment $S$ and create a set of segments $\mathbb{S}_e^\prime$ that includes all labels that are within $t_{low}$ voxels from $e$.
Elements of $\mathbb{S}_e^\prime$ are candidates for merging. 
However, this first pass often leads to too many candidates, requiring an additional pass for further pruning. 
In the second pass, we consider all of the segments in $\mathbb{S}_e^\prime$ for every endpoint $e$. 
If a segment $S^\prime \in \mathbb{S}_e^\prime$ has an endpoint within $t_{high}$ voxels of $e$, the segment $S$ and $S^\prime$ are considered for merging. 
We store the midpoints between the two endpoints as the center of the potential merge in the set $\mathbb{S}_c$. 
This algorithm produces a set of segments to consider for merging. Only these pairs have a corresponding edge in the constructed graph.

@@

We extract a cubic region of interest (ROI) around each endpoint $e$ in $\mathbb{S}_c$ as input to the CNN. 
The CNN receives three input channels for every voxel in the ROI around segments $l_1$ and $l_2$. 
The input in all of the channels is in the range $\{-0.5, 0.5\}$. 
The first channel is $0.5$ only if the corresponding voxel has label $l_1$. 
The second channel is $0.5$ only if the corresponding voxel has label $l_2$. 
The third channel is $0.5$ if the corresponding voxel is either $l_1$ or $l_2$.
We do not use the raw EM image information to avoid the need to retrain the network on datasets that have been stained differently or imaged at different resolution. 
This reduces the need for generating costly manually-labeled ground truth.

\begin{figure}[t]
	\centering
	\includegraphics[width=0.32\linewidth]{./figures/split_error1.png}
	\includegraphics[width=0.32\linewidth]{./figures/split_error2.png}		\includegraphics[width=0.32\linewidth]{./figures/split_error3.png}
	\caption{Three erroneously split segments.}
	\label{fig:merge_candidates}
\end{figure}

@@

@@

\subsubsection{Post-processing}

>>>>

\begin{figure*}[t!]
	\begin{center}
		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/multicut_input.png}}
		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/pre-multicut.png}}
		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/multicut-graph.png}}
		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/post-multicut.png}}
	\end{center}
	\caption{Outline of our approach, from left to right: result of the pixel-based segmentation and agglomeration algorithm; segments of several neurons from the initial segmentation; extracted skeletonized network of those segments; improved 3D reconstruction of the selected segments after graph construction and partitioning with constraints.}
	\label{fig:overview}
\end{figure*}

@@

From the input segmentation we generate a graph $G$ with nodes $N$ and edges $E$ with weights $w_e$. 
The nodes correspond to labeled segments from the input data with edges between merge candidates.
We propose the following three steps to formulate and then partition the graph while obeying constraints from the underlying biology.
First, we only consider merging segments based on a skeletonized representation of the input segmentation.
This is enables us to reduce the number of nodes in the graph based on prior knowledge of the shape of neuronal processes.
Second, we train a convolutional neural network that learns biological constraints based on the shapes of the input segmentation.
This network generates probabilities that segments belong to the same neuron based only on the segmentations. 
An example of one such learned constraint is that neurons have small turning radii (Fig.~\ref{fig:turn-radii}).
Third, we partition the graph using a lifted multicut formulation with additional acyclic constraints to enforce global biological constraints.
The lifted multicut solution is globally consistent which we then augment to produce a tree-structured graph (i.e., one with no cycles).

\subsection{Graph Creation}
%\subsubsection{Node Generation}
\label{sec:skeletonization}

\begin{figure}[t]
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/constraint_error.png}
	\end{minipage}
	\hfill
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/constraint_success.png}
	\end{minipage}
	\caption{Two example candidates for merging. The two segments in the left example do not belong to the same neuron as evidenced by the sharp turning radius.}
	\label{fig:turn-radii}
\end{figure}

\subsubsection{Node Generation}

The simplest node generation strategy creates one node for every unique segment label in the input volume. 
However, some of the labels in the volume correspond to very small structures that are likely the result of segmentation errors, typically in regions with noisy raw image data. 
It is difficult to extract useful shape features from these segments because of their small, often random, shape. 
We prune these nodes from the graph by removing all segments with fewer than a threshold $t_{seg} = 20,000$ voxels. 
This removes on average \TODO{XX}\% of the segments in our  datasets (Sec.~\ref{sec:dataset}), leaving us with around $\TODO{XX}$ nodes per dataset (\TODO{X nodes per micrometer}). 
Despite the large number of segments, these regions only take up \TODO{XX}\% of the total volume on average.

\subsubsection{Edge Generation}

\begin{figure}[t]
	\centering
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/skeleton1.png}		
	\end{minipage}
	\hfill
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/skeleton2.png}		
	\end{minipage}
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/skeleton3.png}		
	\end{minipage}
	\hfill
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/skeleton4.png}		
	\end{minipage}
	\caption{Example skeletons (in black) extracted from segments using a variant of the TEASER algorithm.}
	\label{fig:skeletonization}
\end{figure}

@@

First, we extract a skeleton from each segment in the label volume using a variant~\cite{zhao2014automatic} of the TEASER algorithm~\cite{sato2000teasar}. 
This skeletonization algorithm repeatedly uses Dijkstra's algorithm to find the farthest voxel from a seed location. 
Since this algorithm is non-linear in the number of voxels, we downsample the datasets so that there are voxel samples every $\SI{30}{\nano\meter}$ in each dimension.
Empirically, this reduced the running time for skeletonization by $\TODO{XX\%}$, with minimal reduction in skeleton accuracy ($\TODO{XX\%}$ fewer branches). 
Fig.~\ref{fig:skeletonization} shows four examples of extracted skeletons (in black). 
These skeletons consist of a sequence of \textit{joints}, i.e., locations that are locally a maximum distance from the segment boundary, with line segments connecting successive joints. 
We refer to joints that have only one connected neighbor as \textit{endpoints}. 
We find that approximately \TODO{XX\%} of the segments that are erroneously split have nearby endpoints (Fig.~\ref{fig:merge_candidates}).

Two segments, $s_1$ and $s_2$, receive a corresponding edge if the two following conditions hold.
First, a endpoint in either $s_1$ or $s_2$ is with $t_{low}$ nm of any voxel in the other segment.
Second, there is an endpoint in $s_1$ and an endpoint in $s_2$ that are within $t_{high}$ nm of each other.
We store the midpoints between the two endpoints as the center of the potential merge in the set $\mathbb{S}_c$. 
This algorithm produces a set of segments to consider for merging. 
Only these pairs have a corresponding edge in the constructed graph.

@@

We extract a cubic region of interest (ROI) around each endpoint $e$ in $\mathbb{S}_c$ as input to the CNN. 
The CNN receives three input channels for every voxel in the ROI around segments $l_1$ and $l_2$. 
The input in all of the channels is in the range $\{-0.5, 0.5\}$. 
The first channel is $0.5$ only if the corresponding voxel has label $l_1$. 
The second channel is $0.5$ only if the corresponding voxel has label $l_2$. 
The third channel is $0.5$ if the corresponding voxel is either $l_1$ or $l_2$.
We do not use the raw EM image information to avoid the need to retrain the network on datasets that have been stained differently or imaged at different resolution. 
This reduces the needfor generating costly manually-labeled ground truth.

\begin{figure}[t]
	\centering
	\begin{minipage}{0.32\linewidth}
		\includegraphics[width=\linewidth]{./figures/split_error1.png}		
	\end{minipage}
	\hfill
	\begin{minipage}{0.32\linewidth}
		\includegraphics[width=\linewidth]{./figures/split_error2.png}				
	\end{minipage}
	\hfill
	\begin{minipage}{0.32\linewidth}
		\includegraphics[width=\linewidth]{./figures/split_error3.png}
	\end{minipage}
	\caption{Three erroneously split segments.}
	\label{fig:merge_candidates}
\end{figure}

@@

\subsubsection{Lifted Multicut}

@@

\subsubsection{Topological Constraints}

### 2018-03-13 papers/eccv2018/paper.tex
methods: major updates to skeletonization and neural networks
papers: added much more on biological constraints

\authorrunning{ECCV-18 submission ID \ECCV18SubNumber}
\title{Graph-Based Neural Reconstruction from Skeletonized 3D Networks}

@@

\begin{abstract}
Advancements in electron microscopy image acquisition have created connectomics datasets terabytes in size making manual reconstruction of neuronal structures infeasible.
Current state-of-the-art automatic methods classify neural membranes at the pixel level followed by agglomeration methods to create full neuron reconstructions.
However, these approaches widely neglect global geometric properties that are inherent in the graph structure of neural wiring diagrams.
In this work, we follow pixel-based reconstruction by a global graph-based method to more accurately approximate neural pathways.
We first generate skeletons in 3D from the neuron labels of the pixel-based segmentation.
Next, we simplify this skeletonized 3D network into a graph with nodes corresponding to labels from the segmentation and edges identifying potential locations of split errors.
A CNN classifier produces edge weights for the graph corresponding to the probability of error. 
We formulate the graph into a multicut partitioning problem to improve the final segmentation.
Since the 3D graph is small and encodes top-down information our method is efficient and globally improves the neural reconstruction.
We demonstrate the performance of our approach on multiple real-world connectomics datasets with an average split variation of information improvement of \TODO{XX}\%.
\keywords{connectomics, skeletonization, deep learning}
\end{abstract}

>>>>

\authorrunning{ECCV-18 submission ID \ECCV18SubNumber}
\title{Scalable Graph-Based Neural \\ Reconstruction for Connectomics}

@@

\begin{abstract}
Advancements in electron microscopy image acquisition have created connectomics datasets terabytes in size making manual reconstruction of neuronal structures infeasible.
Current state-of-the-art automatic methods classify neural membranes at the pixel level followed by agglomeration methods to create full neuron reconstructions.
However, these approaches widely neglect global geometric properties and other biological constraints that are inherent in the graph structure of neural wiring diagrams.
In this work, we follow pixel-based reconstruction by a biologically-constrained graph-based method to more accurately approximate neural pathways.
We first generate skeletons in 3D from the neuron labels of the pixel-based segmentation.
Next, we simplify this skeletonized 3D network into a graph with nodes corresponding to labels from the segmentation and edges identifying potential locations of split errors.
A CNN classifier produces edge weights for the graph corresponding to the probability of error. 
We formulate the graph into a lifted multicut partitioning problem to improve the final segmentation.
We demonstrate the performance of our approach on multiple real-world connectomics datasets with an average split variation of information improvement of \TODO{XX}\%.
\keywords{connectomics, skeletonization, deep learning, multicut}
\end{abstract}
